Formalisation of Session-based Lambda Calculus Type Systems

### 2014-11-10 proposal/proposal.tex
Adding some more thoughts to the proposal.

Two aspects of programming language research are relevant for the problem. The first is concerned with static type systems for ensuring runtime safety properties for concurrent communication systems. The second is the increasing use of proof assistants and interactive theorem provers, for studying and mechanically verifying the semantics and properties of programming languages.

@@

Distributed communication systems are becoming increasing prevalent in modern society. To facilitate communication between concurrently executing agents requires complex network protocols to define the type and sequencing of messages. Typically these protocols are implemented in systems programming languages, such as C, which provide no correctness guarantees of the implementation. Furthermore, programming in these languages is often extremely error-prone, as one has a myriad of unsafe features at one's disposal. As more businesses begin to conduct transactions over the Web, it is imperative that they can reason about the communication protocols upon which they rely.

@@

@@

correctness and something what else could we say i think it would appropriate
to mention that the session types studied so far include both pi and lambda
calculi and it may be beneficil if the reader was informed also the pi
calculus paradigms do not fit well into the existing mainstream programming
language paradigms this seems like it should be in the statement of the problem section down below also i should read more of jtod's slides because this writing just keep writing malarky might just be what i need to pound out the pages because it eliminates thoughts as they come into the head what else adopt and focus we have in the problem statement but perhaps those paragraphs should be re-jigged for up here multithreaded functional programming with linear typing for session types a number of these systems use different type systems but they are based on all the lambda calculus fundamentally so to have a generic framework to encode these variants would save on proof engineering effort as is championed by the POPLMARK challenge aimed at reducing the effort for researchers to code programming language metatheory it provides a library needed to perform various things define various parts of programming languages viewed to be difficult to formalise including handling binders and freshness they also mention handling of linear typing but renege on actually formulating this as part of their challenge they insist it is a future work here is the future

@@

@@

@@

>>>>

Two aspects of programming language research are relevant for this project. The first is concerned with static type systems for ensuring runtime safety properties for concurrent communication systems. The second is the increasing use of proof assistants and interactive theorem provers, for studying and mechanically verifying the semantics and properties of programming languages.

@@

Distributed communication systems are becoming increasingly prevalent in modern society. To facilitate communication between concurrently executing agents requires complex network protocols to define the type and sequencing of messages. Typically these protocols are implemented in systems programming languages, such as C, which provide no correctness guarantees of the implementation. Furthermore, programming in these languages is often extremely error-prone, as one has a myriad of unsafe features at one's disposal. As more businesses begin to conduct transactions over the Web, it is imperative that they can reason about the communication protocols upon which they rely.

@@

The increasing use of proof assitants and interactive theorem proving in programming language research has spiked....
POPLMARK
save on proof engineering effort as is championed by the POPLMARK challenge aimed at reducing the effort for researchers to code programming language metatheory it provides a library needed to perform various things define various parts of programming languages viewed to be difficult to formalise including handling binders and freshness they also mention handling of linear typing but renege on actually formulating this as part of their challenge they insist it is a future work here is the future
Metatheory library

@@

correctness and something what else could we say i think it would appropriate
to mention that the session types studied so far include both pi and lambda
calculi and it may be beneficil if the reader was informed also the pi
calculus paradigms do not fit well into the existing mainstream programming
language paradigms this seems like it should be in the statement of the problem section down below also i should read more of jtod's slides because this writing just keep writing malarky might just be what i need to pound out the pages because it eliminates thoughts as they come into the head what else adopt and focus we have in the problem statement but perhaps those paragraphs should be re-jigged for up here multithreaded functional programming with linear typing for session types a number of these systems use different type systems but they are based on all the lambda calculus fundamentally so to have a generic framework to encode these variants would

@@

Roughly can be split into

type systems (pi and lambda)

session type mechanisations

@@

First order approach taken - represent bound variables as de Bruijn indices
In contrast to a higher-order approach
though the first-order encoding has its disadvantages namely one has to reason about variable substitution since without a higher-order approach this is not handled automatically by the meta-language
the locally nameless representation is chosen with a few simplifying assumptions regarding variable substitution and closed terms

Engineering Formal Metatheory
Abstract:

Formalising programming languages within a proof assistant is becoming a
more widely adopted convention within the research community in lieu of hand
written proofs.

Review:

Discussing issues in engineering metatheory for programming languages.
Focus was on the locally nameless representation for binders with names
for free variables and de Brujin indices for bound variables. There was
also novel use of cofinite quantification to rules with bindings to increase
the strength of the induction principles and reduce the amount of renaming
of bound variables (due to freshness issues) significantly. The theory was
tested with formalisations of the simply-typed lambda calculus, System F_sub,
the calculus of constructions. The formalisations followed the POPLMARK
challenge (type soundness, preservation, progress).

@@

The intrinsic encoding as described by~\citeauthor{Benton:2012:STT}~\cite{Benton:2012:STT} is one approach for encoding programming languages based on the lambda calculus. The key idea here is that terms are indexed by their type so as to prevent ill-formed terms from being constructed. However, in the encoding for GV there are issues with using this encoding: $(1)$ it is not immediately clear how to handle linear contexts using the de Bruijn variable encoding, since in the instrinic setting the environment supports weakening; $(2)$ an instrinic encoding of terms would be complicated by the need to enforce session types occur in certain instances (by use of a predicate); and $(3)$ well-typed terms require extra assumptions about binder freshness which cannot reasonably be expressed as a function type. For these reasons, an instrinic approach does not offer much benefit since a well-typed term relation would still need to be defined. I chose the ..... encoding
